---
title: 'Monday Lecture: Iteractive Coding'
author: "Ruth Vergara Reyes"
date: "4/26/2021"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

# **What are we doing today...**

We will be going over loops and the map functions from purrr

# Load Libraries 
```{r}
library(tidyverse)
library(here)
```

# **For Loops  Introduction**

*for loops* are one way to run an iterative process (do something over and over again). However, with the tidyverse you can avoid needing loops for almost any basic coding needs.

```{r}
print(paste("The year is", 2000)) # Simple code


## Putting it in a loop ##
years<-c(2015:2021)
for (i in years){ # set up the for loop where i is the index
  print(paste("The year is", i)) # loop over i
}
```

```{r}
#Pre-allocate space for the for loop
# empty matrix
year_data<-data.frame(matrix(ncol = 2, nrow = length(years)))
# add column names
colnames(year_data)<-c("year", "year_name")

year_data
```


### **Adding to the *for loop* the years **
```{r}
for (i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i]<-paste("The year is", years[i]) # loop over i
}

year_data
```

### **Filling in the year column too**
```{r}
for (i in 1:length(years)){ # set up the for loop where i is the index
  year_data$year_name[i]<-paste("The year is", years[i]) # loop over year name
  year_data$year[i]<-years[i] # loop over year
}

year_data
```

# **Using *for loops* example**

We will be using it to read in multiple .csv files 
```{r}
testdata<-read_csv(here("Week_13", "data", "cond_data","011521_CT316_1pcal.csv"))

glimpse(testdata)
```

### **List files in a directory** 
```{r}
# point to the location on the computer of the folder
CondPath<-here("Week_13", "data", "cond_data")

# list all the files in that path with a specific pattern
# In this case we are looking for everything that has a .csv in the filename

# you can use regex to be more specific if you are looking for certain patterns in filenames
files <- dir(path = CondPath,pattern = ".csv")

files
```


### **Pre-allocate the space for the loop**
```{r}
# make an empty dataframe that has one row for each file and 3 columns
cond_data<-data.frame(matrix(nrow = length(files), ncol = 3))

# give the dataframe column names
colnames(cond_data)<-c("filename","mean_temp", "mean_sal")

cond_data
```


### **Getting the *for loop* ready**
```{r}
raw_data<-read.csv(paste0(CondPath,"/",files[1])) # test by reading in the first [1] file and see if it works; paste0 is used to avoid any white space
head(raw_data)

mean_temp<-mean(raw_data$Temperature, na.rm = TRUE) # calculate a mean of temp column and we use na.rm = true to ignore any NA's in the column
mean_temp
```


### **Filling in the *for loop* **
```{r}
for (i in 1:length(files)){ # loop over 1:3 the number of files
}


for (i in 1:length(files)){ 
raw_data<-read.csv(paste0(CondPath,"/",files[i])) ## adding in the raw data
glimpse(raw_data)
}


for (i in 1:length(files)){  
raw_data<-read.csv(paste0(CondPath,"/",files[i]))
cond_data$filename[i]<-files[i] ## First we want to add the filenames
} 
cond_data

for (i in 1:length(files)){ 
raw_data<-read.csv(paste0(CondPath,"/",files[i]))
cond_data$filename[i]<-files[i]
cond_data$mean_temp[i]<-mean(raw_data$Temperature, na.rm =TRUE) ## adding in the mean for temp.
cond_data$mean_sal[i]<-mean(raw_data$Salinity, na.rm =TRUE) ## adding in the mean for salinity
}

cond_data
```


# **Purrr Introduction**

This is our purrr cheatsheet: 
https://github.com/rstudio/cheatsheets/blob/master/purrr.pdf

### **Map Functions**

The pattern of looping over a vector, doing something to each element and saving the results is so common that the purrr package provides a family of functions to do it for you. There is one function for each type of output:

map() makes a list.
map_lgl() makes a logical vector.
map_int() makes an integer vector.
map_dbl() makes a double vector.
map_chr() makes a character vector.
map_df() makes a dataframe
Each function takes a vector as input, applies a function to each piece, and then returns a new vector that’s the same length (and has the same names) as the input.

We have 3 ways to use the map() function... all lead to the same thing
```{r}
1:10 # Creating a vector from 1 to 10 (we are going to do this 10 times)

1:10 %>%
  map(rnorm, n = 15) # calculate 15 random numbers based on a normal distribution in a list

1:10 %>% 
  map(rnorm, n = 15)  %>% 
  map_dbl(mean) # calculate the mean. It is now a vector which is type "double"
```

### **Now we make our own functions**
```{r}
1:10 %>% # list 1:10
  map(function(x) rnorm(15, x)) %>% # make your own function
  map_dbl(mean)

1:10 %>%
  map(~ rnorm(15, .x)) %>% # changes the arguments inside the function
  map_dbl(mean)
```


### **Bringing in the files using purr instead of a *for loop* **
```{r}
# point to the location on the computer of the folder
CondPath<-here("Week_13", "data", "cond_data")
files <- dir(path = CondPath,pattern = ".csv")
files

files <- dir(path = CondPath,pattern = ".csv", full.names = TRUE)
#save the entire path name
files
```


### **Reading in the files**
```{r}
data<-files %>%
  set_names()%>% # set's the id of each list to the file name
  map_df(read_csv,.id = "filename") # map everything to a dataframe and put the id in a column called filename
data
```


### **Calculate Means**
```{r}
data<-files %>%
  set_names()%>% 
  map_df(read_csv,.id = "filename") %>% 
  group_by(filename) %>%
  summarise(mean_temp = mean(Temperature, na.rm = TRUE),
            mean_sal = mean(Salinity,na.rm = TRUE))
data
```

