---
title: "Troubleshooting Lecture"
author: "Ruth Vergara Reyes"
date: "4/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction:

This lecture will be about how to troubleshoot through R.

# Lecture Notes:

### Errors in Code:

You can tell there is an error from either a:

- red circle with an x in it next to the line number
- red squiggly line under part of your code, indicating a syntax issue.

You can see the error in the console when running code that is broken.

### Messages we can get:
- **Error:** There’s a fatal error in your code that prevented it from being run through successfully. You need to fix it for the code to run.
  - ALWAYS read the message, it can help solve the problem.

- **Warning:** Non-fatal errors (don’t stop the code from running, but this is a potential problem that you should know about).

- **Message:** Here’s some helpful information about the code you just ran (you can hide these if you want to)


# Installing Packages 

```{r}
install.packages('reprex') #reproducible example
install.packages('datapasta') #copy and paste
install.packages('styler') # copy and paste in style
```

### Using the new packages:
```{r}
library(tidyverse)
mpg %>%
  ggplot(aes( x = displ, 
              y = hwy))+
  geom_point(aes(color = class))
```
Using tribble
```{r}
data<-tibble::tribble(
    ~lat,    ~long, ~star_no,
  33.548, -117.805,      10L,
  35.534, -121.083,       1L,
  39.503, -123.743,      25L,
  32.863,  -117.24,      22L,
   33.46, -117.671,       8L,
  33.548, -117.805,       3L
  )

```

