---
title: "Factors Lecture"
author: "Ruth Vergara Reyes"
date: "4/21/2021"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction 
Today we will be learning about factors

# Load Libraries
```{r}
library(tidyverse)
library(here)
library(tidytuesdayR)
```

# Load Data
```{r}
income_mean <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-09/income_mean.csv')
```

# Lecture Notes 

### What is a Factor

It is a specialized version of a character
```{r}
fruits<-factor(c("Apple", "Grape", "Banana"))
fruits
```


##### Factor Booby-traps

NA's are introduced by coercion...
```{r}
test<-c("A", "1", "2")
as.numeric(test)
```

Testing it with factors...
```{r}
test<-factor(test) # covert to factor
as.numeric(test)
```


### {forcats}
```{r}
glimpse(starwars)

starwars %>% 
  filter(!is.na(species))%>% # remove the NAs
  count(species, sort = TRUE)
```

##### Using **fct_lump()**
```{r}
star_counts<-starwars %>%
  filter(!is.na(species)) %>%
  mutate(species = fct_lump(species, n = 3)) %>%
  count(species)

star_counts
```


### Reordering factors: Bar Graphs
```{r}
star_counts %>%
  ggplot(aes(x = species, y = n))+
  geom_col()
```


##### Reording it to be: 

lowest to greatest
```{r}
star_counts %>%
  ggplot(aes(x = fct_reorder(species, n), y = n))+ # reorder the factor of species by n
  geom_col()
```


Greatest to Lowest
```{r}
star_counts %>%
  ggplot(aes(x = fct_reorder(species, n, .desc = TRUE), y = n))+ # reorder the factor of species by n
  geom_col() +
  labs(x = "Species")
```


### Reordering Factors: Line Plots

We will make a plot of the total income by year and quantile across all dollar types.

```{r}
total_income<-income_mean %>%
  group_by(year, income_quintile)%>%
  summarise(income_dollars_sum = sum(income_dollars))%>%
  mutate(income_quintile = factor(income_quintile)) # make it a factor

### Basic Line Plot ###
total_income%>%
ggplot(aes(x = year, y = income_dollars_sum, color = income_quintile))+
  geom_line()

```


##### Using **fct_reorder2**

```{r}
total_income%>%
  ggplot(aes(x = year, y = income_dollars_sum, 
             color = fct_reorder2(income_quintile,year,income_dollars_sum)))+
  geom_line()+
  labs(color = "income quantile")
```


### Reordering Vectors

This is NOT what we want!!!
```{r}
x1 <- factor(c("Jan", "Mar", "Apr", "Dec"))
x1
```

This IS what we want!!
```{r}
x1 <- factor(c("Jan", "Mar", "Apr", "Dec"), levels = c("Jan", "Mar", "Apr", "Dec"))
x1
```

### Subest data with factors

```{r}
starwars_clean<-starwars %>% 
  filter(!is.na(species)) %>% # remove the NAs
  count(species, sort = TRUE) %>%
  mutate(species = factor(species)) %>% # make species a factor
  filter(n>3) # only keep species that have more than 3

starwars_clean

levels(starwars_clean$species)
```


##### Using **fct_drop()**
```{r}
starwars_clean<-starwars %>% 
  filter(!is.na(species)) %>% # remove the NAs
  count(species, sort = TRUE) %>%
  mutate(species = factor(species)) %>% # make species a factor 
  filter(n>3)  %>% # only keep species that have more than 3 
  droplevels() # drop extra levels

levels(starwars_clean$species)
```


### Recode Levels
```{r}
starwars_clean<-starwars %>% 
  filter(!is.na(species)) %>% # remove the NAs
  count(species, sort = TRUE) %>%
  mutate(species = factor(species)) %>% # make species a factor 
  filter(n>3)  %>% # only keep species that have more than 3 
  droplevels() %>% # drop extra levels 
  mutate(species = fct_recode(species, "Humanoid" = "Human"))

starwars_clean
```

