---
title: "Good Plot vs. Bad Plot"
author: "Ruth Vergara Reyes"
date: "3/20/2021"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

# Load Libraries
```{r}
library(tidyverse)
library(here)
library(calecopal)
```


# Load Data
```{r}
CApop <- read_csv(here("good_plot_vs_bad_plot", "data", "CAPopdata.csv"))

view(CApop)
```

# Data Analysis
```{r}
CApop_clean <- CApop %>%
  filter(Population > 444000)
```


# Bad Plot

This plot is a bad plot because of the following reasons:

- It does not explain what the data represents
- It has a distracting background color
- The plot background color makes it difficult to read the names and values
- The color choice makes the title hard to read
- The legend is not necessary and is distracting
- The title can be misleading
  - Someone that is not from California would not know that these are counties in California.

```{r}
ggplot(data = CApop_clean,
       mapping = aes(x = Population,
                     y = County,
                     fill = County))+                          ### Values I used for my axis #
  geom_col()+                                                  ### Type of plot I made #  
  labs(x = "",
       y = "",
       title = "Biggest Counties")+                            ### Labels for my axis #
  theme_classic()+                                             ### Theme for my plot #
  theme(panel.background = element_rect(fill = "hotpink"),
        plot.background = element_rect(fill = "darkblue"),
        title = element_text(size = 15,
                             color = "navyblue"))              ### How I changed different parts of my plot. #

```


# Good Plot

The following is a good plot because:

- The colors that were chosen are not distracting.
- Everything is labeled properly
- The legend that would be distracting is not there
- The plot overall is clean and simple
- Everything that needs labels is labeled
- The title is clear and explains what the data means

```{r, fig.cap="This is a plot that shows the population of the top 19 most populated counties in California", fig.path="../output/"}
ggplot(data = CApop_clean,
       mapping = aes(x = County,
                     y = Population,
                     fill = County))+                     ### How I got my different axis #
  geom_col()+                                             ### Type of plot I made #
  labs(x = "County",
       y = "Total Population",
       title = "Most Populated Counties in California")+  ### How I made the labels #
  theme_classic()+                                           ### The overall theme of the plot #
  scale_x_discrete(guide = guide_axis(angle = 45))+       ### How I got my x axis lables to be angled #
  theme(title = element_text(size = 20,
                             color = "black"),
        axis.title = element_text(size = 13),
        legend.position = "none",
        plot.background = element_rect(fill = "lightgrey"),
        panel.background = element_rect(fill = "white"))                         ### How I changed different aspects of the titles and how I got rid of the unnecessary legend # 
```

